Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-08-18 grove-paper.tex
Light editing pass on sections 2.1 and 2.2.1-2.2.3

\tikzset{graphs/graph style/.style={
 tree layout,
 level sep=1cm,
 nodes={draw, rounded corners=1mm},
 edges={draw, sloped, font={\small}}
}}
\tikzset{graphs/root style/.style={
 %draw=none,
 as={\textbullet$_0$}
}}
%\tikzset{graphs/every edge/.style={}

@@

@@

One is the underlying graph model, which
involves explicit identifiers, graph edges, indexes, and add and delete
operations.

We present this at two levels.
The other is the interface a user might see,
which omits these details unless needed.
Throughout these examples the graph model is presented
as a graph with.

@@

@@

The editor state for each replica in Grove consists of an expression with holes.
For example, \autoref{fig:A simple tree with a hole} shows the simple arithmetic expression~\texttt{x * \hole}, which
has a hole on the right side of the plus operator.

Internally, Grove represents an expression like this as a directed graph with a distinguished root vertex.
Each vertex represents a term and is labeled with a globally unique identifier and a term constructor.
In this figure, \vertex{2} and \vertex{4} have the constructors \texttt{times} and \texttt{var(x)}, respectively.
Each term constructor has an associated set of child positions.
The~\texttt{*} constructor has positions for~\texttt{left} and~\texttt{right} children
and the~\texttt{var} constructors have no child positions but takes an argument naming the particular variable referenced
(e.g.,~\texttt{var(x)} for the variable~\texttt{x}).
The root vertex has a distinguished identifier and a constructor with a single child position.
For simplicity, we visually represent the root with a dot.

For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.

Each edge is labeled with a globally unique identifier (e.g.,~1 and~3 in the figure) and a child position.
An edge indicates that the destination vertex is a child of the origin vertex at the given position.
Visually, we indicate the position of an edge by the location of its origin.
For example, Figure~TODO:1(b) shows the graph corresponding to the term in Figure~TODO:1(a).

TODO: holes

@@

@@

(though as we show in \autoref{sec:formalism} creating a vertex is implicit in edge creation)

@@

TODO: Note that adding an edge implicitly creates.

For the purposes of this paper, we represent identifiers and number literals
indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.

@@

Before \marginpar{DEL?} further exploring how to merge changes from multiple users,
we want to point out a common idiom that we call \emph{wrapping}.
Consider the first graph in \autoref{fig:Wrapping}, which
represents the code~\texttt{$\lambda$ x . x}, and suppose
the user's cursor is in the body field of the lambda (i.e., around the variable
reference to~\texttt{x} in the body of the lambda).

@@

Traditional: no relation (and indent might change)

>>>>

\tikzset{graphs/graph style/.style={
 tree layout,
 level sep=1cm,
 nodes={draw, rounded corners=1mm},
 edges={draw, sloped, font={\small\ttfamily}}
}}
\tikzset{graphs/root style/.style={
 %draw=none,
 as={\textbullet$_0$}
}}
%\tikzset{graphs/every edge/.style={}

@@

TODO: Alice and Bob and narrative continuity

@@

@@

We present this at two levels.
One is the underlying graph model, which
involves explicit identifiers, graph edges, indexes, and add and delete
operations.
The other is the interface a user might see,
which omits these details unless needed.
Throughout these examples the graph model is presented
as a graph with.

@@

The state for each editor instance in Grove consists of an expression containing holes along with a cursor.
For example, \autoref{fig:A simple tree with a hole} shows the arithmetic expression~\texttt{x * \hole}, which
has a hole on the right side of the times operator.

Internally, Grove represents an expression like this as a directed graph with a distinguished root vertex.
For example, \autoref{fig:A simple tree with a hole-b} shows the graph corresponding to the term in \autoref{fig:A simple tree with a hole-a}.
Each vertex represents a term and is labeled with a globally unique identifier and a term constructor.
In \autoref{fig:A simple tree with a hole}, \vertex{2} and \vertex{4} have the constructors \texttt{times} and \texttt{var(x)}, respectively.

Each term constructor has an associated set of child positions.
The~\texttt{*} constructor has positions for~\texttt{left} and~\texttt{right} children
and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).
The root vertex has a distinguished identifier~0 and the constructor \textbullet with the single child position \texttt{root}.

Each edge is labeled with a globally unique identifier (e.g.,~1 and~3 in the figure) and a child position (e.g., \texttt{left} and \texttt{root} in the figure).
An edge indicates that the destination vertex is a child of the origin vertex at the given position.
%Visually, we indicate the position of an edge by the location of its origin.
For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.

Finally, holes are represented by the absence of a child.
For example, in \autoref{fig:A simple tree with a hole-b} the absence of a \texttt{right} edge coming from \vertex{2}
corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.

@@

Note that, for the purposes of this paper, we represent identifiers and number literals
indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.

@@

@@

TODO: (though as we show in \autoref{sec:formalism} creating a vertex is implicit in edge creation)

@@

Next suppose a user inserts cod at a position already occupied by some code.
For example, suppose, the user types~\texttt{+} when the cursor is on \vertex{2}.
If this were a hole, this would simply put a vertex with the~\texttt{+} constructor
as a child of \vertex{0}.

@@

Traditional diff: no relation (and indent might change)
